\label{ch:results_jena}

\subsection{Results}

\subsubsection*{Visitation Rates}			

In total, I made 385 observations, each represents pollinator activity records for 15min in a 80cm x 80cm plot. Accumulated, I analyzed data from 96,25h of observation on 246.4m$^{2}$.

\textit{Onobrychis viciifolia} was the most attractive plant with a maximum of 318 visits in one observation. The per-flower visitation rate (counted visits to all flowers divided through the number of flowers of the focal plant species in the subplot) varied strongly with the attractiveness of the focal species. Per observation, I recorded 1.4 $\pm$ 1.8 ($\pm$ SD) visits per flower with a maximum of 10.7 visits per flower (again \textit{Onobrychis viciifolia}) and 31 observation with no visit at all to the focal species. The per-flower visitation rate was significantly different between the two very attractive species \textit{Geranium pratense} and \textit{Onobrychis viciifolia} and the three less attractive species  \textit{Trifolium pratense}, \textit{Lotus corniculatus} and \textit{Lathyrus pratensis} (\textit{P} $\leq$ 0.001, tab.~\ref{tab:VR_spec}). The subplots contained 3 $\pm$ 1.2 ($\pm$ SD) flowering species including the focal species, the species richness was higher on plot level with 8 $\pm$ 2.4 ($\pm$ SD) flowering species.

\subsubsection*{Frequency Dependence}			
Floral cover and species richness had both individually and in the interaction term with frequency no effect on the visitation rate and were removed from the model (Cover: F$_{df\shorteq1}$ = 1.17, \textit{P} = 0.28; Species Richness: F$ _{df\shorteq1} $ = 1.15, \textit{P} = 0.29). 
%Freq:Cover 0.446 //Freq:SR 0.426

The linear mixed effect model shows an effect of species and frequency individually and with interactions on the per-flower visitation rate (Species: F$_{df\shorteq4}$ = 141.13, \textit{P} $\leq$ 0.0001; Frequency: F$_{df\shorteq1}$ = 18.29, \textit{P} $\leq$ 0.0001; Species x Frequency F$_{ df\shorteq4}$ = 5.2, \textit{P} $\leq$ 0.001, tab.~\ref{tab:anova}). Interestingly, frequency contributes also as quadratic and cubic term with its interactions to species explanatory power to the model, giving the relationship a non-linear character (Tab.~\ref{tab:anova}). Figure~\ref{fig:LME} shows the third degree polynomial relationship of all focal species and the summed data except \textit{Lathyrus pratensis} which has a quadratic relationship. The sigmoid curve is defined by a strong increase for frequencies below 20\% followed by a minimum between 50 and 80\% depending on the species before raising again with increasing dominance of the focal species. However, the visitation rate of \textit{Lathyrus pratensis} presents a maximum at 60\% frequency and decreases afterwards. 




